Timing and magnitude of climate driven range shifts in transboundary fish stocks challenge their management

### 2021-11-25 Scripts/Initial_Analysis.Rmd
Updated stats results with tresh_two

```{r emergence_map_plot, eval = F}
# -------------------------- #
# Call plot for both tresholds
# -------------------------- #
tresh_list <- c("tresh_one","tresh_two")
for(i in 1:2){

@@

@@

# Number of coastal nations
Eme_Res %>% 
  rename(geo_region = un_sub_region) %>% 
  group_by(eezid,sau_fishing_entity,geo_region,tresh,taxon_key) %>% 
  summarise(mean_yr = mean(emerging_yr,na.rm=T)) %>% 
  right_join(top_stocks_per_country,
             by = c("sau_fishing_entity","geo_region","tresh","taxon_key")
  ) %>% 
  filter(mean_yr <= 2021) %>% 
  group_by(variable,tresh) %>% 
  summarise(
    n_countries = length(unique(sau_fishing_entity))
  ) %>% 
  mutate(per = n_countries/length(unique(Sau_data$fishing_entity_id)))

# Number of shifting stocks per geo-region
Eme_Res %>% 
  rename(geo_region = un_sub_region) %>% 
  group_by(eezid,sau_fishing_entity,geo_region,tresh,taxon_key) %>% 
  summarise(mean_yr = mean(emerging_yr,na.rm=T)) %>% 
  right_join(top_stocks_per_country,
             by = c("sau_fishing_entity","geo_region","tresh","taxon_key")
  ) %>% 
  filter(mean_yr <= 2021) %>% 
  group_by(geo_region) %>% 
  tally() %>% 
  arrange(n) %>% 
  top_frac(0.25,n) # top 25%

@@

hist(Stats_data$emerging_yr) # Positively skewed
qqnorm(Stats_data$emerging_yr) # not normally distributed
skewness(Stats_data$emerging_yr) #1.02
kurtosis(Stats_data$emerging_yr) #2.82

@@

Reg_kw
# Kruskal-Wallis rank sum test
# 
# data:  un_sub_region by emerging_yr
# Kruskal-Wallis chi-squared = 287.23, df = 93, p-value < 2.2e-16

kruskalmc(emerging_yr~un_sub_region, data = Reg_diff_data) %>% 
  as.data.frame() %>% 
  rownames_to_column("comparrison") %>% 
  # head()
  View()
write_csv("kw_regional.csv")

>>>>

```{r emergence_map_plot, eval = F}
# -------------------------- #
# Call plot for both tresholds
# -------------------------- #
tresh_list <- c("tresh_one","tresh_two")
i = 2
for(i in 2){

@@

  ggsave("mean_boxplot.png",boxplot,width = 10)

@@

# Number of coastal nations
Eme_Res %>% 
  rename(geo_region = un_sub_region) %>% 
  group_by(eezid,sau_fishing_entity,geo_region,tresh,taxon_key) %>% 
  summarise(mean_yr = mean(emerging_yr,na.rm=T)) %>% 
  right_join(top_stocks_per_country,
             by = c("sau_fishing_entity","geo_region","tresh","taxon_key")
  ) %>% 
  # View()
  filter(mean_yr <= 2021) %>% 
  group_by(variable,tresh) %>% 
  # View()
  summarise(
    n_countries = length(unique(sau_fishing_entity))
  ) %>% 
  mutate(per = n_countries/length(unique(Sau_data$fishing_entity_id))*100)

# Number of shifting stocks per geo-region
Eme_Res %>% 
  rename(geo_region = un_sub_region) %>% 
  group_by(eezid,sau_fishing_entity,geo_region,tresh,taxon_key) %>% 
  summarise(mean_yr = mean(emerging_yr,na.rm=T)) %>% 
  right_join(top_stocks_per_country,
             by = c("sau_fishing_entity","geo_region","tresh","taxon_key")
  ) %>% 
  filter(mean_yr <= 2021,
         variable == "value",
         tresh == "tresh_two",
         !is.na(geo_region)) %>% 
  group_by(geo_region,tresh) %>% 
  tally() %>% 
  arrange(desc(n)) %>% 
  # pull(n) %>% 
  # sum()

139/263

@@

hist(Stats_data$emerging_yr) # Positively skewed
qqnorm(Stats_data$emerging_yr) # not normally distributed
skewness(Stats_data$emerging_yr) #0.61
kurtosis(Stats_data$emerging_yr) #2.11

@@

Reg_kw
# Kruskal-Wallis rank sum test
# 
# data:  un_sub_region by emerging_yr
# Kruskal-Wallis chi-squared = 242.11, df = 93, p-value = 3.128e-15

kruskalmc(emerging_yr~un_sub_region, data = Reg_diff_data) %>% 
  as.data.frame() %>% 
  rownames_to_column("comparrison") %>% 
  # head()
  # View()
write_csv("kw_regional.csv")

### 2021-11-25 Scripts/Manuscript.Rmd
Updated stats results with tresh_two

Most fisheries management is poorly adapted to address species’ shifting distributions because of changing ocean conditions. The effectiveness of fisheries management for transboundary species is challenged by shifts in stocks’ distribution under climate change [@Pinsky:2018cb; @Pinsky:2014by; @Pinsky:2020jx; @Oremus:2020id]. In most cases, catch or fishing effort quotas for transboundary stocks are based on historical records and do not necessarily consider the full distribution range of the stock [@FredstonHermann:2018kp], nor the effects of a changing climate on fish stocks and associated fisheries [@PalaciosAbrantes:2020tv; @Sumby:2021gw]. While many transboundary stocks experience natural seasonal variation in their distribution (e.g., migration, reproductive cycle), misalignment between fisheries resources' allocation and distributional shifts beyond such natural variation have previously resulted in unsustainable fishing levels and international disputes [@Miller:2013iv; @Spijkers:2017ij; @Crespo:2020cs]. For example, Pacific salmon species (*Oncorhynchus sp.*), jointly managed by Canada and the U.S., were heavily fished in the early 1990s after a stock shift favored the Alaskan fleet over the Canadian fleet [@Miller:2013iv]. Continuing climate change is expected to exacerbate such patterns [@Pinsky:2018cb; @Sumaila:2011bq]. Understanding when climate change will affect the sharing dynamics of transboundary stocks and the intensity of the resulting impacts is important for developing climate resilient international ocean governance and achieving the goals set out under the 2030 Agenda [@Pinsky:2018cb; @Link:2010ei; @Sumaila:2020tp].

Here, we employ a mechanistic population dynamic model driven by output from a comprehensive Earth system model with ten ensemble members to project transboundary stocks’ distribution across the 280 EEZs of 198 coastal countries/political entities under a high greenhouse gas emissions scenario (Methods). **We define a ‘stock’ unit as a species shared by neighboring EEZs.** Furthermore, we  developed a Transboundary Shift Index to evaluate range shifts in the shared distribution of transboundary stocks. The index is based on changes in the distance of the stock’s abundance centroid and that of the EEZs that share a given stock (Fig. \@ref(fig:toeFigure1N)). By employing the concept of *time of emergence*, we estimate the year in which the 633 transboundary species (9,132 stocks) are projected to shift their shared distribution beyond historical variations (Fig. 1). Such estimation is based on two thresholds representing a 67 and a 95% probability of shift. Finally, we apply the Game Theory concept of *threat point* [See @Sumaila:2020tp; @Munro:1979dg for its application to fisheries] to quantify the intensity of changes in the shared distribution of transboundary stocks between neighboring EEZs (Methods). This assumes that countries are adapted to the historical sharing, but no country would be willing to collaborate if the future catch proportion of the transboundary stock is lower than it has ever been [@PalaciosAbrantes:2020vt; @Sumaila:2020tp]. We treat each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a managed population [@PalaciosAbrantes:2020tv] and only consider shared stocks between neighboring EEZs (i.e., excluding the high seas).

@@

The analyses detailed herein are based on 633 exploited marine transboundary species that account for 80% of the catch taken from the world’s EEZs between 2005 and 2014 [@PalaciosAbrantes:2020tv]. **We define a (transboundary) "stock" unit when a species was shared by a pair of neighboring EEZs [@PalaciosAbrantes:2020tv; @Teh:2015gd], resulting in a total of 9,132 transboundary stocks.** Using the transboundary species Atlantic cod (*Gadus morhua*) as an example and this study’s definition of transboundary stock, the United States and Canada share a stock of Atlantic cod, and Canada and Greenland share a separate stock of Atlantic cod, but the United States and Greenland do not as they do not have juxtaposing EEZs. We defined the boundaries of the world’s EEZs using the Sea Around Us spatial division (updated 1 July 2015; http://www.seaaroundus.org), noting that it subdivides the EEZs of 198 coastal states into 280 regions (Fig. \@ref(fig:toeFigureS1)) including island territories. We determined the intersections between polygons using the R package sf [@PackagesfSimpleF:2018vp]. Each EEZ was categorized by geopolitical region according to the United Nations (https://population.un.org/wpp/DefinitionOfRegions/) and biome [@Reygondeau:2019uh]. The habitat preference of each species was determined following FishBase (http://www.fishbase.org) for fish species and SeaLifeBase (http://www.fishbase.org) for invertebrates (Table \@ref(tab:toeTableS1)). For each stock and EEZ, we used Sea Around Us data to estimate catch and fishing revenue from fishing activities. We report both average catch and revenue for the last available decade (2005-2014) [@Zeller2016; @Tai2017; @Sumaila:2015uc].

@@

**While the total number of shifting stocks is important for fisheries management, having high-valuable stocks shifting sooner could be a much bigger concern than having many low-value stocks shifting. Thus, we looked at the year of range shift of the top five most valuable transboundary stocks of each country (New Fig 4). On average, countries are projected to see a range shift of the top five most valuable stocks by 2038 $\pm$ 25 years under threshold two (2028 $\pm$ 22 years under threshold one). Moreover, our projections suggest that these high value stocks are already shifting in 52 (27%) coastal nations under threshold two (n = 85 44% under threshold one), with the regions of Latin America and the Caribbean (n = 63), Sub-Saharan Africa (n = 59), South-eastern Asia  (n = 28), and Polynesia (n = 19) presenting the larger number of shifting stocks (25% of total high value stocks).**

@@

These uncertainties also affect the projection of species distributions [@Lotze:2019ce; @BryndumBuchholz:2019kh]. Overall, multiple upper trophic level models present overall agreement in terms of directional change of fish biomass, but variability in the magnitude of that change [@Lotze:2019ce]. Further research that reduces uncertainty in the NPP response at the base of ocean food webs, alongside large ensemble simulations of multiple species distribution models could lead to smaller structural uncertainty of fish and fisheries models [@BryndumBuchholz:2019kh]. **Another important source of uncertainty in this study is the utilization of political boundaries (EEZ) to define stocks, rather than biologically-defined populations. While this method might define some stocks that do not necessarily align with biologically-defined sub-populations within an EEZ, in many EEZs, fisheries are often managed at the species level [@MAP:2017uf] and sub-populations are potentially interconnected [@Ramesh:2019va], thus, providing additional ecological ground for our analysis [@Dunn:2019ca]. Moreover, not limiting the species to, at least, an EEZ could result in an overestimation of our results (NOTE: See Ovando *et al.* https://www.pnas.org/content/118/23/e2025958118).** Reproducing our analysis regionally, where spatially explicit stock data is available would allow to generate better constrained results and potentially identify different types of shared stock shifts at meta-population level [@Link:2010ei; @Archambault:2016ie]. Addressing these uncertainties systematically can serve as a roadmap for future studies to provide additional information to inform policy towards sustainable and equitable international fisheries management under climate change.

>>>>

Most fisheries management is poorly adapted to address species’ shifting distributions because of changing ocean conditions. The effectiveness of fisheries management for transboundary species is challenged by shifts in stocks’ distribution under climate change [@Pinsky:2018cb; @Pinsky:2014by; @Pinsky:2020jx; @Oremus:2020id]. In most cases, catch or fishing effort quotas for transboundary stocks are based on historical records and do not necessarily consider the full distribution range of the stock [@FredstonHermann:2018kp], nor the effects of a changing climate on fish stocks and associated fisheries [@PalaciosAbrantes:2020tv; @Sumby:2021gw]. While many transboundary stocks experience natural seasonal variation in their distribution (e.g., migration, reproductive cycle), misalignment between fisheries resources' allocation and distributional shifts beyond such natural variation have previously resulted in unsustainable fishing levels and international disputes [@Miller:2013iv; @Spijkers:2017ij; @Crespo:2020cs]. For example, Pacific salmon species (*Oncorhynchus sp.*), jointly managed by Canada and the U.S., were heavily fished in the early 1990s after a stock shift favored the Alaskan fleet over the Canadian fleet [@Miller:2013iv]. Continuing climate change is expected to exacerbate such patterns [@Pinsky:2018cb; @Sumaila:2011bq]. **Recent studies have looked at global gains (Pinsky et. al., 2018) and losses (Oremus et al., 2020) of transboundary stocks within the world’s EEZs as well as regional sharing dynamics between neighbouring countries (Palacios-Abrantes et al., 2020, Sumaila et al., 2020) towards the end of the century because of climate change. However, previous studies fail to identify changes in the distribution of current transboundary stocks at the global level including the actual timeline of shifts in shared stocks, nor the intensity of such shifts.** Understanding when climate change will affect the sharing dynamics of transboundary stocks and the intensity of the resulting impacts is important for developing climate resilient international ocean governance and achieving the goals set out under the 2030 Agenda [@Pinsky:2018cb; @Link:2010ei; @Sumaila:2020tp].

Here, we employ a mechanistic population dynamic model driven by output from a comprehensive Earth system model with ten ensemble members to project transboundary stocks’ distribution across the 280 EEZs of 198 coastal countries/political entities under a high greenhouse gas emissions scenario (Methods). **We adopt the definition of a ‘stock’ unit developed and applied in Palacios-Abrantes *et al.* 2020; Teh & Sumaila, 2015, which considers a stock unit as a species shared by neighbouring EEZs.** Furthermore, we  developed a Transboundary Shift Index to evaluate range shifts in the shared distribution of transboundary stocks. The index is based on changes in the distance of the stock’s abundance centroid and that of the EEZs that share a given stock (Fig. \@ref(fig:toeFigure1N)). **We base our analysis on the 633 transboundary species identified by Palacios-Abrantes (2020) defining a ‘stock’ unit as each time a transboundary species is shared by neighbouring EEZs. By employing the concept of time of emergence, we estimate the year in which the stocks are projected to shift their shared distribution beyond historical variations (Fig. \@ref(fig:toeFigure1N)).** Such estimation is based on two thresholds representing a 67 and a 95% probability of shift. Finally, we apply the Game Theory concept of *threat point* [See @Sumaila:2020tp; @Munro:1979dg for its application to fisheries] to quantify the intensity of changes in the shared distribution of transboundary stocks between neighboring EEZs (Methods). This assumes that countries are adapted to the historical sharing, but no country would be willing to collaborate if the future catch proportion of the transboundary stock is lower than it has ever been [@PalaciosAbrantes:2020vt; @Sumaila:2020tp]. **We treat each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a managed population [@PalaciosAbrantes:2020tv] and only consider shared stocks between neighboring EEZs (i.e., excluding the high seas).**

@@

The analyses detailed herein are based on 633 exploited marine transboundary species that account for 80% of the catch taken from the world’s EEZs between 2005 and 2014 [@PalaciosAbrantes:2020tv]. **We define a (transboundary) "stock" unit when a species was shared by a pair of neighboring EEZs [@PalaciosAbrantes:2020tv; @Teh:2015gd], resulting in a total of 9,132 transboundary stocks.** **Let us consider the transboundary species Atlantic cod (Gadus morhua) as an example. Atlantic cod’s distribution in the Northwest Atlantic ranges from the United States to Greenland. Based on the previous definition of a transboundary stock, the United States and Canada share a stock of Atlantic cod, and Canada and Greenland share a separate stock of Atlantic cod, but the United States and Greenland do not share a stock as they do not have juxtaposing EEZs.** We defined the boundaries of the world’s EEZs using the Sea Around Us spatial division (updated 1 July 2015; http://www.seaaroundus.org), noting that it subdivides the EEZs of 198 coastal states into 280 regions (Fig. \@ref(fig:toeFigureS1)) including island territories. We determined the intersections between polygons using the R package sf [@PackagesfSimpleF:2018vp]. Each EEZ was categorized by geopolitical region according to the United Nations (https://population.un.org/wpp/DefinitionOfRegions/) and biome [@Reygondeau:2019uh]. The habitat preference of each species was determined following FishBase (http://www.fishbase.org) for fish species and SeaLifeBase (http://www.fishbase.org) for invertebrates (Table \@ref(tab:toeTableS1)). For each stock and EEZ, we used Sea Around Us data to estimate catch and fishing revenue from fishing activities. We report both average catch and revenue for the last available decade (2005-2014) [@Zeller2016; @Tai2017; @Sumaila:2015uc].

@@

**While the total number of shifting stocks is important for fisheries management, having more valuable stocks shifting sooner could be a much bigger concern than having many low-value stocks shifting. Thus, we looked at the year of range shift of the top five most valuable transboundary stocks of each country (Fig. \@ref(fig:toeFigure3)). On average, countries are projected to see a range shift of the top five most valuable stocks by 2038 $\pm$ 25 years under threshold two (2028 $\pm$ 22 years under threshold one). Moreover, our projections suggest that these high value stocks are already shifting in 129 (66%) coastal nations under threshold two (n = 154, 79% under threshold one), with the regions of Latin America and the Caribbean (n = 62), Sub-Saharan Africa (n = 36), Polynesia (n = 22), and Southern Europe (n = 19) presenting the larger number of high value shifting stocks (50% of shifting high value stocks).**

@@

These uncertainties also affect the projection of species distributions [@Lotze:2019ce; @BryndumBuchholz:2019kh]. Overall, multiple upper trophic level models present overall agreement in terms of directional change of fish biomass, but variability in the magnitude of that change [@Lotze:2019ce]. Further research that reduces uncertainty in the NPP response at the base of ocean food webs, alongside large ensemble simulations of multiple species distribution models could lead to smaller structural uncertainty of fish and fisheries models [@BryndumBuchholz:2019kh]. **Another important source of uncertainty in this study is the utilization of political boundaries (EEZ) to define stocks, rather than biologically-defined populations. While this method might define some stocks that do not necessarily align with biologically-defined sub-populations within an EEZ, in many EEZs, fisheries are often managed at the species level [@MAP:2017uf] and sub-populations are potentially interconnected [@Ramesh:2019va], thus, providing additional ecological ground for our analysis [@Dunn:2019ca]. Moreover, the population structure of most exploited species in the world are poorly- or un- defined [@ref]. Future studies that focus on particular shared stocks could use more specific unit stock definition that is available for the cases.** Reproducing our analysis regionally, where spatially explicit stock data is available would allow to generate better constrained results and potentially identify different types of shared stock shifts at meta-population level [@Link:2010ei; @Archambault:2016ie]. Addressing these uncertainties systematically can serve as a roadmap for future studies to provide additional information to inform policy towards sustainable and equitable international fisheries management under climate change.
